The relative roles of genetic, ecological and environmental factors in species radiations

### 2018-02-20 paper/draft.tex
add a lot of refs

One of the first and to this day highly influential formal theories of island biodiversity is MacArthur and Wilson's seminal equilibrium theory of island biogeography (ETIB). %TODO: ref
In it, species richness on islands is the result of opposing rates of immigration and extinction, both of which in turn are modified by island size (area) and its distance from the mainland (isolation).
This relationship was since successfully applied to a multitude of islands %TODO: refs
, but also island like systems, such as coral reefs, lakes/ponds or forest fragments. %TODO: refs
This general applicability, however, cannot hide the fact that the ETIB fails to deliver a mechanistic explanation of the processes at play.
It has particularly been criticized for its lack of considering speciation. %TODO: refs

About a decade ago, island biogeography theory gained new momentum by Whittaker et al.'s General Dynamic Model (GDM). %TODO: ref
The conceptual framework traces island communities along the geomorphological trajectory of hotspot islands,
whith islands emerging from sea level, increasing over time in size and topographic complexity, before eroding and subsiding in the ocean again.
In doing so, the model incorporates concepts of niche theory, speciation and species succession.
Yet, there are still many processes and drivers deemed important, which are not considered by any formal theory and often little investigated.

@@

In a recent review \cite{leidinger2017biodiversity}, we analyzed a powerful alternative to field experiments in island biogeography: mechanistic (/simulation) models.
Although many of the reviewed models integrate multiple processes at a time, unfortunately, most models failed to include more explicit genetic or micro evolution processes.
Instead, some models used concepts such as ``protracted speciaton'' \cite{rosindellXXX},
where speciation occurs with a delay from the actual
divergence event if the population remains isolated during that time.
But nature is usually more complex than this (cf.\ Herben et al.\ 2005) and genetic mechanisms can have important ramifications for speciation/radiation and species richness.

@@

For example, the rapid switching of pollinator syndromes in species of the genus \textit{Petunia} %use package for this!
seems to be a consequence of the tight genetic linkage between loci for pollination syndrome traits (Hermann et al. 2013) % cite properly!

>>>>

One of the first and to this day highly influential formal theories of island biodiversity is MacArthur and Wilson's seminal equilibrium theory of island biogeography (ETIB). %TODO: ref
In it, species richness on islands is the result of opposing rates of immigration and extinction, both of which in turn are modified by island size (area) and its distance from the mainland (isolation).
This relationship was since successfully applied to a multitude of islands (Lomolina \& Weiser 2001, Triantis et al.\ 2011)%TODO: refs
, but also island like systems, such as coral reefs, lakes/ponds or forest fragments (Kelly et al.\ 1989, Peay et al.\ 2007, ). %TODO: refs
This general applicability (Lomolino 2000), however, cannot hide the fact that the ETIB fails to deliver a mechanistic explanation of the processes at play.
It has particularly been criticized for its lack of considering speciation (see e.g.\ Losos \& Schluter 2000). %TODO: refs

About a decade ago, island biogeography theory gained new momentum by Whittaker et al.'s General Dynamic Model (GDM, Whittaker et al.\ 2007, 2008). %TODO: ref
The conceptual framework traces island communities along the geomorphological trajectory of hotspot islands,
whith islands emerging from sea level, increasing over time in size and topographic complexity, before eroding and subsiding in the ocean again.
In doing so, the model incorporates concepts of niche theory, speciation and species succession.
Yet, there are still many processes and drivers deemed important, which are not considered by any formal theory and often little investigated.

@@

In a recent review \cite{leidinger2017biodiversity}, we analyzed a powerful alternative to field experiments in island biogeography: mechanistic (/simulation) models.
Although many of the reviewed models integrate multiple processes at a time, unfortunately, most models failed to include more explicit genetic or micro evolution processes.
Instead, some models used concepts such as ``protracted speciaton'' \cite{rosindellXXX},
where speciation occurs with a delay from the actual
divergence event if the population remains isolated during that time.
But nature is usually more complex than this (cf.\ Silvertown 2004, but see Herben et al.\ 2005) and genetic mechanisms can have important ramifications for speciation/radiation and species richness.

@@

For example, the rapid switching of pollinator syndromes in species of the genus \textit{Petunia} %use package for this!
seems to be a consequence of the tight genetic linkage between loci for pollination syndrome traits (Hermann et al.\ 2013) % cite properly!
